Contracts for Contracts

### 2023-05-09 paper/introduction.tex
more intro

We observe that popular smart contract languages (such as Solidity) do
not provide \emph{rich, effective, convenient} means to specify and enforce
contract behaviors.
Simple first-order assertion primitives (e.g., @require@) are indeed common,
but they cannot effectively check callable addresses or function values, which
are a common pattern used in smart contract programs.
Moreover, many of the critical behavior in smart contracts are \emph{temporal}.
For instance, the defense to the DAO attack \note{cite} is to
prevent ``reentrancy'' via external functions, which requires that the function
of interest cannot be invoked \emph{before} the current call of this function is finished.

@@

%Apart from the investigating effort to certify and verify smart contract
%programs, language designers are responsible to empower programmers with better
%linguistic solutions to secure smart contracts in the first place.
In this paper, we develop practical \emph{behavioral
software contracts} that provide a lightweight solution for securing
\emph{smart contracts}. In particular, we develop a specification language and
monitor for higher-order and temporal behaviors in the Solidity language.

\subsection*{\textbf{Behavioral Contracts for Smart Contracts}}

Behavioral contracts provide a mechanism for programmers to specify assumptions
and guarantees of software components. The specified assumptions and guarantees
will be monitored and reported when violated.
Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
Design-by-Contract methodology \cite{DBLP:conf/tools/Meyer98a}, behavioral
contracts have been widely adopted in Java, Haskell, C\#, Python, Racket,
Elixir \cite{DBLP:conf/erlang/0001BBHMEF22}, etc. \todo{cite}

\todo{why BC is good here?}

At the conceptual level, behavioral contracts is complementary to smart
contracts, in the sense that they are both agreement -- between software
components for the former, and between transactional parties for the later.
Thus, the two terms ``contract'' indeed mean the same thing but just different
levels of enforcement.

In this work, we design and implement a practical extension of Solidity, the
most popular language for building smart contract programs. The extension
enriches the Solidity language with \emph{higher-order} and \emph{temporal}
behavioral contracts, which are absent in existing approaches.
\todo{allows furthermore downstream tasks, fuzzing, static verification,
	natural language to formal spec etc.}

@@

Higher-order contracts provides a way to specify and enforce pre- and
post-conditions of untrusted functions. \todo{...}

@@

We develop a core model of the extension, including its syntax,
operational semantics, and discuss its soundness guarantee.
With this extension of higher-order temporal behavioral contracts, we examine
several cases of reported bugs and vulnerabilities and show that
our approach is effective in preventing these attacks.
\zz{Let's mention the amount of funds we could save if we can prevent these attacks.}
We also evaluate the overhead (gas consumption) of our approach
compared to manually inserted assertion checks, showing that \todo{...}

>>>>

We observe that popular smart contract languages (such as Solidity) do
not provide \emph{rich, effective, convenient} means to specify and enforce
contract behaviors.
Simple first-order assertion primitives (e.g., @require@) are indeed common,
but they cannot effectively check callable addresses or function values, which
are a common pattern used in smart contract programs.
Moreover, many of the critical condition in smart contracts are \emph{temporal} --
the order of happened events matters.
For instance, the defense to the DAO attack \note{cite} is to
prevent ``reentrancy'' via external functions, which requires that the function
of interest cannot be invoked \emph{before} the current call of this function is finished.

@@

%Apart from the investigating effort to certify and verify smart contract
%programs, language designers are responsible to empower programmers with better
%linguistic solutions to secure smart contracts in the first place.
In this paper, we develop practical \emph{behavioral
software contracts} that provide a lightweight solution to secure
\emph{smart contracts}.
In particular, we develop a specification language and monitor for higher-order
and temporal behaviors in the Solidity language, demonstrating that they
are an effective approach to

\subsection*{\textbf{Behavioral Contracts for Solidty}}

Similar to smart contracts, behavioral contracts is an agreement too.  They
specify the assumptions and guarantees between software components.
In languages with behavioral contracts, the specified assumptions and
guarantees will be monitored and reported when violated.
Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
Design-by-Contract methodology \cite{DBLP:conf/tools/Meyer98a}, behavioral
contracts have been widely adopted in Java, Haskell, C\#, Python, Racket,
Elixir \cite{DBLP:conf/erlang/0001BBHMEF22}, etc. \note{cite}

Solidity is a programming language to build smart contract programs that run on
the Ethereum platform.
As one of the most popular languages for smart contracts, Solidity has been
used \todo{to do cool things}.
However, Solidity and its running platform Ethereum is a complex system.
Carelessly written contracts can cause serious bugs and loses.
We identify \todo{solidity's tricky feature}

To build robust smart contract programs, we design and implement a practical
extension of Solidity, namely \lang.
The extension enriches the Solidity language with \emph{higher-order} and
\emph{temporal} behavioral contracts, which are absent in existing approaches.
With higher-order and temporal behavioral specification, many downstream
tasks become further amenable, including but not limited to static verification,
guided testing, or specification extraction.

@@

In Solidity, callbacks can be in two forms, addresses and higher-order functions.

@@

\paragraph{}

We develop a core model of \lang, including its syntax, operational semantics,
and discuss its soundness guarantee.
With this extension of higher-order temporal behavioral contracts, we examine
several cases of reported bugs and vulnerabilities and show that
our approach is effective in preventing these attacks.
\zz{Let's mention the amount of funds we could save if we can prevent these attacks.}
We also evaluate the overhead (gas consumption) of our approach
compared to manually inserted assertion checks, showing that \todo{...}
